Introduction_outline

### 2019-01-25 Write_up/Writing/Introduction_outline.Rmd
references

As previously mentioned, upwelling winds in EBUSs are generated by pressure differences between continental thermal lows and oceanic highs. As warming intensifies, pressure gradients between land and sea increases, resulting in a greater intensity of upwelling winds [@bakun2010]. In rare cases, projected upwelling increases may overcome the countervailing effects of upper-ocean warming and stratification to cause regional cooling [@auad]. Research studies reveal declines in ocean oxygen levels as a result of ocean warming and reduced ventilation from stratification and circulation changes (Keeling et al. 2010). Similarly, low oxygen levels are also possible because of shifts in the composition of sinking organic material [@hofmann; @doney].

@@

Understanding the relationship between global warming and climate change as a result of anthropogenic activity is a vast and challenging process. In a bid to further this understanding, Rockström et al. (2009) introduced the concept of planetary boundaries. According to this concept, within each of these boundaries exists a level of safety that should not be exceeded. For example, the planetary boundary for biodiversity loss, has been surpassed (Steffen et al., 2015). Similarly, the safety boundary for warming of the globe by greenhouse gas emission has also been surpassed. The idea of planetary boundaries is essential in that it helps our understanding of the threats we cause via the continued destruction of the natural environment. Consequently, studies of anthropogenic influences on the ecosystem are generally divided into land or sea. Terrestrial ecosystems are more familiar to the general populace; whereas marine ecosystems are associated with high levels of primary production (Chavez et al., 2011) and are highly at risk as a result of climate change (Österblom et al., 2017).  This has been proven to decrease primary production thus affecting the trophic levels and leading to trophic turbulence and destruction of marine fauna (Lewandowska et al., 2014; Dell et al., 2014, McCauley et al., 2015).

Climate change similarly affects coastal species such as oysters, corals and, sea grass. Lord (2017) also found evidence that suggests that a change in climate as a result of rising sea temperatures may increase the spread of invasive species and increase the chances of an outbreak of diseases (Altizer et al., 2013). Research by Mead et al (2013) found that coastal ecosystems are at high risk as a result of climate change, later, Whitfield et al (2016) confirmed that coastal bio-diversity have been observed to be affected. Given the importance of primary productivity, it was determined that the Benguela upwelling systerm within the South African region should become the focus of this thesis. Here the increase in global temperatures are one of the main focus points along with the chlorophyll concentration. By making use of these two measurements, I am able to analyses how chlorophyll concentration vary with a change in temperature.

@@

Benguela Niños are defined as anomalous warm events occurring between the southward flowing Angola current and the Benguela upwelling system off south western Africa (Shannon et al. 1986). These warm events profoundly impact on fisheries and the climate of this region. They are proven to induce unexpected rainfall events and drastically influence fish abundance and distribution (Rouault et al. 2003, Boyer et al. 2001). Evidently, Benguela Niños are expressed as regions of abnormal, persistent high sea surface temperatures. Many researchers often associate the equatorial interannual variability pattern in the Atlantic to the ENSO phenomenon found within the Pacific region (Sutton et al. 2000; Zebiak 1993). Benguela Niños have been recorded to occur in 1934, 1949, 1963, 1984 and 1995 (Shannon et al. 1986; Gammelsrød et al. 1998). Picaut (1985) suggested that eastern tropic oceans are governed by remote wind stress effects through equatorial wave dynamics. Anomalies in trade winds may result in Kelvin waves that propagate eastward along the equator, inducing a deepening or a lifting of the thermocline. It is possible that equatorial Kelvin waves may play a role in the evolution of Benguela Niños.

@@

>>>>

As previously mentioned, upwelling winds in EBUSs are generated by pressure differences between continental thermal lows and oceanic highs. As warming intensifies, pressure gradients between land and sea increases, resulting in a greater intensity of upwelling winds [@bakun2010]. In rare cases, projected upwelling increases may overcome the countervailing effects of upper-ocean warming and stratification to cause regional cooling [@auad]. Research studies reveal declines in ocean oxygen levels as a result of ocean warming and reduced ventilation from stratification and circulation changes [@Keeling]. Similarly, low oxygen levels are also possible because of shifts in the composition of sinking organic material [@hofmann; @doney].

@@

Understanding the relationship between global warming and climate change as a result of anthropogenic activity is a vast and challenging process. In a bid to further this understanding, Rockström et al. (2009) introduced the concept of planetary boundaries. According to this concept, within each of these boundaries exists a level of safety that should not be exceeded. For example, the planetary boundary for biodiversity loss, has been surpassed [@steffen]. Similarly, the safety boundary for warming of the globe by greenhouse gas emission has also been surpassed. The idea of planetary boundaries is essential in that it helps our understanding of the threats we cause via the continued destruction of the natural environment. Consequently, studies of anthropogenic influences on the ecosystem are generally divided into land or sea. Terrestrial ecosystems are more familiar to the general populace; whereas marine ecosystems are associated with high levels of primary production [@Chavez] and are highly at risk as a result of climate change [@osterblom]. This has been proven to decrease primary production thus affecting the trophic levels and leading to trophic turbulence and destruction of marine fauna [@lewandowska; @dell, @mcCauley].

Climate change similarly affects coastal species such as oysters, corals and, sea grass. Lord (2017) also found evidence that suggests that a change in climate as a result of rising sea temperatures may increase the spread of invasive species and increase the chances of an outbreak of diseases [@altizer]. Research by Mead et al (2013) found that coastal ecosystems are at high risk as a result of climate change, later, Whitfield et al (2016) confirmed that coastal bio-diversity have been observed to be affected. Given the importance of primary productivity, it was determined that the Benguela upwelling systerm within the South African region should become the focus of this thesis. Here the increase in global temperatures are one of the main focus points along with the chlorophyll concentration. By making use of these two measurements, I am able to analyses how chlorophyll concentration vary with a change in temperature.

@@

Benguela Niños are defined as anomalous warm events occurring between the southward flowing Angola current and the Benguela upwelling system off south western Africa [@shannon1986]. These warm events profoundly impact on fisheries and the climate of this region. They are proven to induce unexpected rainfall events and drastically influence fish abundance and distribution [@rouault; @boyer]. Evidently, Benguela Niños are expressed as regions of abnormal, persistent high sea surface temperatures. Many researchers often associate the equatorial interannual variability pattern in the Atlantic to the ENSO phenomenon found within the Pacific region (Sutton et al. 2000; Zebiak 1993). Benguela Niños have been recorded to occur in 1934, 1949, 1963, 1984 and 1995 [@shannon; @gammelsrod]. Picaut (1985) suggested that eastern tropic oceans are governed by remote wind stress effects through equatorial wave dynamics. Anomalies in trade winds may result in Kelvin waves that propagate eastward along the equator, inducing a deepening or a lifting of the thermocline. It is possible that equatorial Kelvin waves may play a role in the evolution of Benguela Niños.

@@

Picaut, J., 1985: Major dynamics affecting the eastern tropical At-
lantic and Pacific Oceans. CalCOFI Report, Vol 26.
